---
title       : Predicting Battlefield 4 Conquest Games
subtitle    : How using historical data can help improve the player experiance
author      : Miroslaw Horbal
job         : 
framework   : html5slides   # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]     # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Predicting Battlefield 4 Conquest Games

Why should we try to predict the outcome of games? 

1. We develop statistical models of factors that are important in determining a winning team
2. We can learn about natural imbalances in the game from historical data
3. We can develop automated server-balancing tools via discrete optimization to 
improve the player experiance

---

## Factors that determine a winning team
- Is it really possible to predict the outcome of a game of conquest using 
historical data? 
- Are there factors that exist that can be used to predict the winning team?

```{r LoadData, echo=FALSE, cache=TRUE}
playerData <- read.csv("../PeerAssignment/data/playerStats_v1.csv")
reportData <- read.csv("data/reportData.csv.bz2")
```

We think: Yes!

- We collected the in-game statistics of `r nrow(playerData)` players on PS3 and 
Xbox 360 and `r nrow(reportData)` Conquest battle reports
- We computed the average difference of kill/death ratio (KDR) between both teams
- We can predict the outcome of a game of Conquest with 
`r with(reportData, round(mean((winner=="Team1")==(kdr>0))*100, 2))`% accuracy using _only_
KDR

---

```{r PlotKDRDensity, echo=FALSE, cache=TRUE, fig.align='center', fig.height=5, fig.width=8, warning=FALSE}
library(ggplot2)
plt <- qplot(kdr, data=reportData, geom="density", fill=winner, xlim=c(-1.5,1.5)) + 
  geom_density(alpha=0.1) + ggtitle("Average difference in KDR between Team 1 and Team 2") +
  xlab("Avg Kill/Death Ratio (KDR) difference")
print(plt)
```

It's clear from this figure that we determine factors that can be used for predicting
the outcome of games. KDR is only one of _many_ factors that can be used, and we can expect
improvments in prediction accuracy as we include other correlated factors

---

## Discovering imbalances in games
- We can use historical report data to explore natural imbalances in game design
- By comparing the average number of wins across maps for each team we can estimate balance flaws in
map design

```{r GetMapImbalance, echo=FALSE, cache=TRUE, fig.align='center', fig.height=5, fig.width=9, warning=FALSE}
library(reshape2)
tbl <- prop.table(with(reportData, table(map,winner)),1)
tbl <- melt(tbl)
plt <- ggplot(tbl, aes(map, value)) + geom_bar(aes(fill=winner), stat="identity") + 
  theme(axis.text.x=element_text(angle=90, size=12)) + xlab("Map") + ylab("Win Frequency") +
  geom_hline(aes(yintercept=0.5), linetype="dashed", size=1) + ggtitle("Map vs Win Frequency")
print(plt)
```

---

## Imrpoving automated team balancing 

- We can use report data and player data to improve automated server balancing so games are not one-sided
- We can train a probabalistic model for $h(t_1, t_2) = P(t_1 \text{Wins} | t_1, t_2 )$ where $t_1, t_2$ are
players for Team 1 and 2 respectively 
- Using discrete optimization algorithms such as branch and bound we can efficiently optimize the player 
grouping of Team 1 and Team 2 to solve: $$\arg\!\min_{t_1,t_2} |h(t_1, t_2)-0.5|$$
- This optimization can naturally include squad constraints